---
title: "RegressionFinalProject"
author: "Stephanie"
date: "10/5/2021"
output: html_document
---

```{r}
library(haven)
library(broom)
library(GGally)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(car)
```

```{r}
df <- read_sav("ElPasoHispanics2012_DJ_040521.sav")
```


```{r}
df %>% 
  select(# mental_score
         MentalHealthMedCLN, DepressionCLN, AnxietyCLN, Q66.1, Q66.2, Q66.3, 
         # health_score
         DiabetesMedCLN, SeizureCLN, BloodCLN, SkinCLN, SleepCLN, VitaminCLN, LiverCLN, ThyroidCLN, HeadCLN, HeartCLN, DentalCLN, FluColdCLN, CholestorolCLN, CancerCLN, BurnsCLN, InfectionCLN, GastritisStomCLN, AsthmaCLN, ArthritisCLN, AnemiaCLN, AllergyCLN, ADHDCLN, HeartburnCLN, HiBloodPresCLN, PhysInjuryCLN, 
         Age, Gender, HousingCLN, Citizenship, GenCLN, HighestEducation, YearsEdu, FBFriendsCLN, 
         #insurance
         Q63A, 
         YearsEPCLN, SESCLN, 
         # smoke
         Q83B, Q83bb,  
         # drink
         Q84a, Q84b, 
         Q85, Q85a, Q85redcln, 
         Q88, 
         # drug_score
         Q90.1A, Q90.2A, Q90.3A, Q90.4A, Q90.5A, Q90.6A, 
         Q95A,  
         Q96ACLN, 
         # medical debts
         Q118B, 
         # how many sleep
         Q139A, 
         #weights
         Weightscl
         ) ->
  pj_df
```

```{r}
# write it to a smaller csv
#write.csv(pj_df, "regression_fp_data.csv", row.names = FALSE)
```

```{r}
pj_df <- read_csv("regression_fp_data.csv")
#glimpse(pj_df)
```

# Y Variable: Health Score

```{r}
# Create variable: health_score
pj_df$health_score <- apply(pj_df[1:31], 1, function(x) length(which(x==1)))
distinct(unique(pj_df[c("health_score")]))

# Create variable: drug_score
pj_df$drug_score <- apply(pj_df[43:48], 1, function(x) length(which(x==1)))
distinct(unique(pj_df[c("drug_score")]))
```

```{r}
# check na values in each column
pj_df %>% 
  summarise(across(everything(), ~ sum(is.na(.x))))
```

```{r}
# replace certain number which indicates na values without "Age" column
pj_df[1:31, 33:60][pj_df[1:31, 33:60] == c(41,78,99)] <- NA 

pj_df %>% 
  summarise(across(everything(), ~ sum(is.na(.x))))
```

```{r}
# Filter only needed column
pj_df %>% 
  select(-(1:31), -(43:48)) ->
  pj_df
```

# Replace NA to mean value of column "YearsEPCLN" (have 662 NAs, others have less than 10)
```{r}
# check how many NAs are in each column
pj_df %>% 
  summarise(across(everything(), ~ sum(is.na(.x))))

# df YearsEPCLN without NA
pj_df %>% 
  select(YearsEPCLN) ->
  ins_wo_na

na.omit(ins_wo_na) -> ins_wo_na

ins_wo_na[ins_wo_na == "Inf"] <- NA
ins_wo_na[ins_wo_na == "-Inf"] <- NA

ins_wo_na %>% 
  drop_na ->
  ins_wo_na

# replace NAs with mean
round(mean(ins_wo_na$YearsEPCLN), 0) # mean: 21
pj_df$YearsEPCLN[is.na(pj_df$YearsEPCLN)] <- 21
```

# Overall F-test: p-value < 0.05 indicates that at least one have relationship with y (health_score)
```{r}
# With the variable we think will have relationship
lm_begin <- lm(health_score ~ Age + Gender + HousingCLN + Citizenship + GenCLN + HighestEducation + YearsEdu + FBFriendsCLN + Q63A + YearsEPCLN + SESCLN + Q85redcln + drug_score + Q118B + Q139A, data = pj_df)
glance(lm_begin)
```

```{r, message=FALSE}
pj_df %>% 
  select(health_score, Gender, HousingCLN, Citizenship, GenCLN, YearsEdu) -> for_gg_df1
ggpairs(for_gg_df1)

pj_df %>% 
  select(health_score, FBFriendsCLN, Q63A,YearsEPCLN, SESCLN, drug_score, Q139A) -> for_gg_df2
ggpairs(for_gg_df2)
```


# Test SLR in Big Model
1. With Age
```{r}
lm_out1 <- lm(health_score ~ Age, data = pj_df, weights = Weightscl)
tidy(lm_out1)
```

2. With Gender
```{r}
lm_out2 <- lm(health_score ~ Gender, data = pj_df, weights = Weightscl)
tidy(lm_out2)

# plot
qplot(x = Gender, y = health_score, data = pj_df)
qplot(x = Gender, y = health_score, data = pj_df) + geom_jitter()
#residual plot
aout1 <- augment(x = lm_out2)
qplot(x = .fitted, y = .resid, data = aout1) + geom_jitter()
```

3. With HousingCLN
```{r}
lm_out3 <- lm(health_score ~ factor(HousingCLN), data = pj_df, weights = Weightscl)
tidy(lm_out3)
```

4. With Citizenship
```{r}
lm_out4 <- lm(health_score ~ factor(Citizenship), data = pj_df, weights = Weightscl)
tidy(lm_out4)
```

5. With GenCLN
```{r}
lm_out5 <- lm(health_score ~ factor(GenCLN), data = pj_df, weights = Weightscl)
tidy(lm_out5)

# plot
#qplot(x = factor(GenCLN), y = health_score, data = pj_df)
#residual plot
#lm_cat <- lm(health_score ~ GenCLN, data = pj_df, weights = Weightscl)
#aout2 <- augment(x = lm_cat)
#qplot(x = .fitted, y = .resid, data = aout2)
```

6. With HighestEducation
```{r}
lm_out6 <- lm(health_score ~ factor(HighestEducation), data = pj_df, weights = Weightscl)
tidy(lm_out6)
```

7. With YearsEdu
```{r}
lm_out7 <- lm(health_score ~ YearsEdu, data = pj_df, weights = Weightscl)
tidy(lm_out7)

# plot
#qplot(x = YearsEdu, y = health_score, data = pj_df)
#residual plot
#aout3 <- augment(x = lm_out7)
#qplot(x = .fitted, y = .resid, data = aout3)
#qplot(x = .fitted, y = .resid, data = aout3) + geom_jitter()
```

8. With FBFriendsCLN
```{r}
lm_out8 <- lm(health_score ~ FBFriendsCLN, data = pj_df, weights = Weightscl)
tidy(lm_out8)
```

9. With Insurance
```{r}
lm_test1 <- lm(health_score ~ Q63A + YearsEPCLN + factor(SESCLN), data = pj_df, weights = Weightscl)
tidy(lm_test1)

# Q63A(Insurance) is relevant
# plot
qplot(x = Q63A, y = health_score, data = pj_df) + geom_jitter()
#residual plot
lm_63A <- lm(health_score ~ Q63A, data = pj_df, weights = Weightscl)
tidy(lm_63A)
aout4 <- augment(x = lm_63A)
qplot(x = .fitted, y = .resid, data = aout4) + geom_jitter()

# YearsEPCLN is relevant
# plot
qplot(x = YearsEPCLN, y = health_score, data = pj_df) + geom_jitter()
#residual plot
lm_yearsep <- lm(health_score ~ YearsEPCLN, data = pj_df, weights = Weightscl)
tidy(lm_yearsep)
aout5 <- augment(x = lm_yearsep)
qplot(x = .fitted, y = .resid, data = aout5)
```

10. With Drinking
```{r}
lm_test2 <- lm(health_score ~ Q85redcln, data = pj_df, weights = Weightscl)
tidy(lm_test2)

# plot
qplot(x = Q85redcln, y = health_score, data = pj_df) + geom_jitter()
#residual plot
aout6 <- augment(x = lm_test2)
qplot(x = .fitted, y = .resid, data = aout6)
```

11. With Drug Score
```{r, message=FALSE}
lm_test6 <- lm(health_score ~ drug_score, data = pj_df, weights = Weightscl)
tidy(lm_test6)

# plot
qplot(x = drug_score, y = health_score, data = pj_df) + geom_jitter() # Unequal Variance?
#residual plot
aout7 <- augment(x = lm_test4)
qplot(x = .fitted, y = .resid, data = aout7) + geom_jitter() + geom_smooth()

# try log(y)
pj_df %>% 
  mutate(log_health_score = log(health_score)) ->
  try_log

na.omit(try_log) -> try_log

try_log[try_log == "Inf"] <- NA
try_log[try_log == "-Inf"] <- NA

try_log %>% 
  drop_na ->
  try_log

lm_log_drug <- lm(log_health_score ~ drug_score, data = try_log, weights = Weightscl)
tidy(lm_log_drug)
aout_log1 <- augment(x = lm_log_drug)
qplot(x = .fitted, y = .resid, data = aout_log1) # not better

# try log both x and y
pj_df %>% 
  mutate(log_health_score = log(health_score), 
         log_drug_score = log(drug_score)) ->
  try_log2

na.omit(try_log2) -> try_log2

try_log2[try_log2 == "Inf"] <- NA
try_log2[try_log2 == "-Inf"] <- NA

try_log2 %>% 
  drop_na ->
  try_log2

lm_log_both <- lm(log_health_score ~ log_drug_score, data = try_log2, weights = Weightscl)
tidy(lm_log_both)
aout_log2 <- augment(x = lm_log_both)
qplot(x = .fitted, y = .resid, data = aout_log2) # residual plot
```

12. With Medical Debts
```{r}
lm_test7 <- lm(health_score ~ Q118B, data = pj_df, weights = Weightscl)
tidy(lm_test7)
```

13. With Sleep
```{r}
lm_test8 <- lm(health_score ~ Q139A, data = pj_df, weights = Weightscl)
tidy(lm_test8)

# plot
qplot(x = Q139A, y = health_score, data = pj_df) + geom_jitter()
#residual plot
aout8 <- augment(x = lm_test8)
qplot(x = .fitted, y = .resid, data = aout8)

# log(x)
pj_df %>% 
  mutate(log_Q139A = log(Q139A)) ->
  try_log

na.omit(try_log) -> try_log

try_log[try_log == "Inf"] <- NA
try_log[try_log == "-Inf"] <- NA

try_log %>% 
  drop_na ->
  try_log

all(is.na(try_log)) # check any na left

#try_log$log_Q139A <- as.numeric(gsub("\\.", "", try_log$log_Q139A))
#try_log$health_score <- as.numeric(gsub("\\.", "", try_log$health_score))

lm_log_Q139A <- lm(health_score ~ log_Q139A, data = try_log, weights = Weightscl)
tidy(lm_log_Q139A)
aout_log <- augment(x = lm_log_Q139A)
qplot(x = .fitted, y = .resid, data = aout_log)

# log not working, we should just remove outliers! More than 20 hrs sleep doesn's make sense!
pj_df %>% 
  filter(Q139A < 20) ->
  try_rm_olrs

lm_rm_o <- lm(health_score ~ Q139A, data = try_rm_olrs, weights = Weightscl)

qplot(x = Q139A, y = health_score, data = try_rm_olrs) + geom_jitter() # plot

aout_rm <- augment(x = lm_rm_o)
qplot(x = .fitted, y = .resid, data = aout_rm) #residual plot
```

# MLR Model
```{r}
# Filter only needed model
pj_df %>% 
  select(health_score, Gender, Q63A, YearsEPCLN, drug_score, Q139A, Weightscl) ->
  final_df

final_df %>% 
  rename("Health_Score" = "health_score", 
         "Insurance" = "Q63A",
         "Drug_Score" = "drug_score", 
         "Sleep" = "Q139A") -> 
  final_df

final_df %>% 
  filter(Sleep < 20) ->
  final_df

# fit the MLR
lm_out <- lm(Health_Score ~ Gender + Insurance + YearsEPCLN + Drug_Score + Sleep, data = final_df, weights = Weightscl)
tidy(lm_out)



# Type I sum of squares
anova(lm_out)
# we can get: 
SSR(X1|X2, X3, X4, X5) = 6694
SSR(X2|X1, X3, X4, X5) = 1977
SSR(X3|X1, X2, X4, X5) = 4182
SSR(X4|X1, X2, X3, X5) = 2204
SSR(X5|X1, X2, X3, X4) = 4455
# conclusion: Adding Gender decrease the sum of squares by 6694, but adding in Insurance only decreased it by additional 1977, while every other predictors present

Anova(lm_out)
```

```{r, message=FALSE}
# GGPairs
final_df %>% 
  select(Health_Score, Gender, Insurance, YearsEPCLN, Drug_Score, Sleep) ->
  pairs1
ggpairs(data = pairs1)
```

# Residual PLot

```{r}
# MLR
lm_rich <- lm(Health_Score ~ Gender + Insurance + YearsEPCLN + Drug_Score + Sleep, data = final_df, weights = Weightscl)
tidy(lm_rich)
summary(lm_rich)

a_rich <- augment(lm_rich)
qplot(x = .fitted, y = .resid, data = a_rich)
```

```{r}
# Check if there is Multicollinearity
vif(lm_rich)

# GGpairs
final_df %>% 
  select(Gender, Insurance, YearsEPCLN, Drug_Score) ->
  pairs2
ggpairs(data = pairs2)

# Run LR
lm_check <- lm(Health_Score ~ Gender + Insurance + YearsEPCLN + Drug_Score, data = final_df, weights = Weightscl)
tidy(lm_check)
  
lm_check1 <- lm(Health_Score ~ Gender, data = final_df, weights = Weightscl)
lm_check2 <- lm(Health_Score ~ Insurance, data = final_df, weights = Weightscl)
lm_check3 <- lm(Health_Score ~ YearsEPCLN, data = final_df, weights = Weightscl)
lm_check4 <- lm(Health_Score ~ Drug_Score, data = final_df, weights = Weightscl)

tidy(lm_check1)
tidy(lm_check2)
tidy(lm_check3)
tidy(lm_check4)

# Gender and Q63A has slightly multicollinearity
lm_a <- lm(Health_Score ~ Gender + Insurance, data = final_df, weights = Weightscl)
tidy(lm_a)

# Try test the r-squared
glance(lm(Gender ~ Sleep, data = final_df, weights = Weightscl))$r.squared
```

```{r}
# Fit Categorical Variable (Gender) in
final_df %>% 
  mutate(Gender = as.character(Gender)) ->
  test_df

# convert to factor
test_df %>% 
  mutate(Gender = parse_factor(Gender, levels = c("1", "2"))) ->
  test_df

# fit
lm_gen <- lm(Health_Score ~ Gender + Insurance + YearsEPCLN + Drug_Score + Sleep, data = test_df, weights = Weightscl)
tidy(lm_gen, conf.int = T) # reference level: Gender2
```

